RegressionPvNP

### 2017-04-27 Patient Report.Rmd
Moved pre-processing pipeline code

patient.info = read.patient.info(patient.file)
patient.info$Patient = gsub("/", "_", patient.info$Patient)
head(patient.info)

@@

@@

save(patient.data, file='patients.Rdata')

## Note the number of probes per patient/sample
```
\pagebreak

# Data

@@

1. Segmented values 
```{r}
head(patient.data[['AD0258']]$seg.vals)
```

```{r echo=F, warning=F, message=F, fig.height=10, fig.width=10}
ggplot( melt(patient.data[['AD0258']]$seg.vals[,-(1:5)]), aes(x=value) ) + 
        facet_wrap(~variable,scales = "free_x") + 
        geom_histogram( fill=I("blue"),col=I("lightblue"), binwidth=0.03) + labs(title='AD0258 segmented value distribution', x='', y='')
```

2. Row (e.g. per-region across timepoints) normalized values from the matrix above
```{r}
head(patient.data[['AD0258']]$norm.seg.vals)
```

```{r echo=F, warning=F, message=F, fig.height=10, fig.width=10}
ggplot( melt(patient.data[['AD0258']]$norm.seg.vals), aes(x=value) ) + 
        facet_wrap(~variable,scales = "free_x") + 
        geom_histogram( fill=I("blue"),col=I("lightblue"), binwidth=0.3) + labs(title='AD0258 normalized segmented value distribution', x='', y='')

@@

  do.call('grid.arrange', list(g1,g2, top=paste(name, status, sep=" - ")))
})

@@

prog.regions = sapply(var.seg.lengths[sum.patient.data$Patient], nrow)

@@

chrs = read.table('hg19_chrom.txt',sep='\t', header=T)
var.seg.lengths = lapply(var.seg.lengths, function(x) merge(x, chrs, by.x='chr', by.y='chr') )

segment.lengths= lapply(segment.lengths, function(x) merge(x, chrs, by.x='chr', by.y='chr') )

opcov = sapply(segment.lengths[prog.pt], function(x) {
  sum(as.numeric(x[['seg.len']]))/sum(as.numeric(chrs$length))
})

onpcov = sapply(segment.lengths[np.pt], function(x) {
  sum(as.numeric(x[['seg.len']]))/sum(as.numeric(chrs$length))
})

chr.ratios = lapply(var.seg.lengths, function(x) x[['seg.len']]/x[['length']])
# Segment coverage, variable regions
ttR = t.test(unlist(chr.ratios[subset(sum.patient.data, Status == 'P')$Patient]), 
             unlist(chr.ratios[subset(sum.patient.data, Status == 'NP')$Patient]))

@@

pcov = median(sapply(var.seg.lengths[prog.pt], function(x) 
  sum(as.numeric(x[['seg.lengths']]))/sum(as.numeric(x[['length']])) ), na.rm=T)

@@

## Grouping the chromosomes by length 1-12, 13-22
#t.test(subset(np.regions, chrom %in% c(1:12))$chr.ratio, subset(np.regions, chrom %in% c(13:22))$chr.ratio)

#t.test(subset(prog.regions, chrom %in% c(1:12))$chr.ratio, subset(prog.regions, chrom %in% c(13:22))$chr.ratio)

@@

So it appears that adjusting for chromosome length (looking at the ratio of the chromosome covered by segments rather than the raw bp) a significant difference in the coverage of the genome by CN segments that are variable between progressors and non: t-test p-value `r signif(ttR$p.value, 2)`.  Overall bp coverage of the sWGS is `r round(mean(c(opcov, onpcov)), 2)*100`%  (P `r round(mean(opcov),2)`, NP `r round(mean(onpcov), 2)`).

@@

gene.cnt = arrange(gene.cnt, -Progressors)
gene.p = subset(gene.cnt, Progressors >= Non.Progressors)

@@

p = ggplot(m, aes(y=Gene.Symbols, x=variable)) + 
  geom_tile(aes(fill=value), color='white') + 
  scale_fill_gradient('n patients',low='white',high='royalblue4') + labs(x="",y='CGC Genes')
p + geom_point(data=m[m$CN == 'gain' & m$value > 0,], aes(y=Gene.Symbols, x=variable), size=2, shape=24, fill='red', color='red') +
  geom_point(data=m[m$CN == 'loss' & m$value > 0,], aes(y=Gene.Symbols, x=variable), color='green4', size=2, shape=25, fill='green4')

#gc = gene.cnt %>% group_by(Gene.Symbols) %>% summarise(Progressors=sum(Progressors), Non.Progressors=sum(Non.Progressors))
#gc = arrange(subset(gc, Progressors >= Non.Progressors), -Progressors)
#m = melt(gc, id.vars=c('Gene.Symbols'))
#m$Gene.Symbols = factor(m$Gene.Symbols, levels= m$Gene.Symbols[c(nrow(m):1)] ) # order
#ggplot(m, aes(y=Gene.Symbols, x=variable)) + 
#  geom_tile(aes(fill=value), color='white') + 
#  scale_fill_gradient('n patients',low='white',high='royalblue4') + labs(x="",y='CGC Genes')
```

>>>>

dataset = 'Training'
patient.info = read.patient.info(patient.file, set=dataset)
patient.info$Patient = gsub("/", "_", patient.info$Patient)
head(patient.info)

@@

# Description of '`r dataset`' data set

@@

save(patient.data, file=paste(dataset, '_patients.Rdata', sep=''))

example.pt = subset(sum.patient.data, Status == 'P' & total.samples > 2)$Patient[1]
## Note the number of probes per patient/sample
```
\pagebreak

@@

1. Segmented values (progressor): `r example.pt` 
```{r}
head(patient.data[[example.pt]]$seg.vals)
```

```{r echo=F, warning=F, message=F, fig.height=10, fig.width=10}
ggplot( melt(patient.data[[example.pt]]$seg.vals[,-(1:5)]), aes(x=value) ) + 
        facet_wrap(~variable,scales = "free_x") + 
        geom_histogram( fill=I("blue"),col=I("lightblue"), binwidth=0.03) + labs(title=paste(example.pt,' segmented value distribution'), x='', y='')
```

2. Row (e.g. per-region across timepoints) normalized values from the matrix above
```{r}
head(patient.data[[example.pt]]$norm.seg.vals)
```

```{r echo=F, warning=F, message=F, fig.height=10, fig.width=10}
ggplot( melt(patient.data[[example.pt]]$norm.seg.vals), aes(x=value) ) + 
        facet_wrap(~variable,scales = "free_x") + 
        geom_histogram( fill=I("blue"),col=I("lightblue"), binwidth=0.3) + labs(title=paste(example.pt,' normalized segmented value distribution'), x='', y='')

@@

  do.call('grid.arrange', list(g1,g2, top=paste(name, status, sep=" - "), ncol=2))
})

@@

n.var.regions = sapply(var.seg.lengths, nrow)
wilcox.test(n.var.regions[prog.pt], n.var.regions[np.pt])

@@

chrs = read.table('hg19_chrom.txt',sep='\t', header=T)

# Segment lengths, variable regions
var.seg.lengths = lapply(var.seg.lengths, function(x) merge(x, chrs, by.x='chr', by.y='chr') )

# All segments
segment.lengths = lapply(segment.lengths, function(x) merge(x, chrs, by.x='chr', by.y='chr') )

# Genome coverage
genome.coverage = sapply(segment.lengths, function(x) {
  sum(as.numeric(x[['seg.len']]))/sum(as.numeric(chrs$length))
})

# No difference in coverage
covtt = wilcox.test(genome.coverage[prog.pt], genome.coverage[np.pt])

# Variable region length by ratio of chromosome
chr.ratios = lapply(var.seg.lengths, function(x) x[['seg.len']]/x[['length']])

melted = melt(chr.ratios)
melted = merge(sum.patient.data[,c('Patient', 'Status')], melted, by.x='Patient', by.y='L1')

chr.ratio.tt = t.test(unlist(chr.ratios[prog.pt]), unlist(chr.ratios[np.pt]))

ggplot(melted, aes(Status, value, fill=Status)) + geom_boxplot() + 
  labs(title=paste("Variable region length, p-value", ifelse(chr.ratio.tt$p.value < 0.001, "<0.001", signif(chr.ratio.tt$p.value,2))),y='Segment Length/Chr length')

@@

# Genome coverage of the variable regions
pcov = median(sapply(var.seg.lengths[prog.pt], function(x) 
  sum(as.numeric(x[['seg.lengths']]))/sum(as.numeric(x[['length']])) ), na.rm=T)

@@

## Grouping the chromosomes by length 1-12, 13-22
get.chrs<-function(x, chrs) { subset(x, chr %in% chrs)$chr.ratio }

long.chr.tt = t.test(unlist(lapply(var.seg.lengths[prog.pt], get.chrs, chrs=c(1:12))), unlist(lapply(var.seg.lengths[np.pt], get.chrs, chrs=c(1:12))))
short.chr.tt = t.test(unlist(lapply(var.seg.lengths[prog.pt], get.chrs, chrs=c(13:22))), unlist(lapply(var.seg.lengths[np.pt], get.chrs, chrs=c(13:22))))

@@

So it appears that adjusting for chromosome length (looking at the ratio of the chromosome covered by segments rather than the raw bp) a significant difference in the coverage of the genome by CN segments that are variable between progressors and non: t-test p-value `r signif(chr.ratio.tt$p.value, 2)`.  Overall bp coverage of the sWGS is `r round(mean(genome.coverage), 2)*100`%  (P `r round(mean(genome.coverage[prog.pt]),2)`, NP `r round(mean(genome.coverage[np.pt]), 2)`).

@@

gene.cnt = arrange(gene.cnt, -Progressors)
## RAD51B is screwing up the plot simply because it shows 16 non-progressors (8 progressors) with a gain but the next closest is 9 so we'll just filter that out for now
gene.cnt = subset(gene.cnt, gene.cnt$Gene.Symbols != 'RAD51B')

#gene.p = subset(gene.cnt, Progressors >= Non.Progressors)
gene.p = gene.cnt

@@

ggplot(m, aes(y=Gene.Symbols, x=variable)) + 
  geom_tile(aes(fill=value), color='white') + 
  scale_fill_gradient('n patients',low='white',high='red') + labs(x="",y='CGC Genes')
# + geom_point(data=m[m$CN == 'gain' & m$value > 0,], aes(y=Gene.Symbols, x=variable), size=2, shape=24, fill='red', color='red') +
#  geom_point(data=m[m$CN == 'loss' & m$value > 0,], aes(y=Gene.Symbols, x=variable), color='green4', size=2, shape=25, fill='green4')

gc = gene.cnt %>% group_by(Gene.Symbols) %>% summarise(Progressors=sum(Progressors), Non.Progressors=sum(Non.Progressors))
gc = arrange(subset(gc, Progressors >= Non.Progressors), -Progressors)
gc = gc[1:60,]
m = melt(gc, id.vars=c('Gene.Symbols'))
#m$value = as.factor(m$value)
m$Gene.Symbols = factor(m$Gene.Symbols, levels= m$Gene.Symbols[c(nrow(m):1)] ) # order
ggplot(m, aes(y=Gene.Symbols, x=variable)) + 
  geom_tile(aes(fill=value), color='white') + 
  scale_fill_gradient('n patients',low='white',high='red') + labs(x="",y='CGC Genes') + ggtitle("Top 50 genes only")

### 2017-04-27 RegressionPvNP.Rmd
Moved pre-processing pipeline code

sum.patient.data = summarise.patient.info(patient.info)
sum.patient.data = as.data.frame(subset(sum.patient.data, Patient %in% names(patient.data))) ## For now
pander(sum.patient.data[,c('Patient', 'Status', 'start.year', 'end.year','total.samples','highest.path')])

@@

```{r xvalpt, message=F, warning=F, echo=F, fig.height=6, fig.width=6}
precisionRecall<-function(actual, predicted) {
    retrieved <- sum(predicted==1)
    recall <- sum(predicted==1 & actual==1) / sum(actual==1)
    precision <- sum(predicted==1 & actual==1) / retrieved
    f1 = signif(2*(1/(1/precision+1/recall)), 2)
    return(data.frame('precision'=precision, 'recall'=recall, 'F1'=signif(f1, 2)))
  }

@@

cv.patient.glmnet<-function(x,y,lambda,patients) {
  message(paste("Running", length(unique(patients$group)), "folds"))

  cv.class = matrix(nrow=5, ncol=length(lambda))
  auc = matrix(nrow=5, ncol=length(lambda))

  fit.e = list()
  for (i in 1:length(unique(patients$group))) {  ## This is not stable. 
    message(paste(i, "fold"))
    test.rows = which(rownames(x) %in% subset(patients, group == i)$Samplename)
    test = x[test.rows,]
    training = x[-test.rows,]
    # pre-spec lambda seq
    fit <- glmnet(training, y[-test.rows], lambda=lambda, family='binomial', alpha=1) 
    # autoplot(fit) + theme(legend.position="none")

    pred.class <- predict(fit, test, type='class') # check this later
    auc[i,] = apply(pred.class, 2, function(pd) {
      if (sum(as.numeric(pd)) <= 0) return(0)
      #as.numeric(pROC::auc(pROC::roc(response=as.numeric(pd), predictor=labels[test.rows])))  
      precisionRecall(labels[test.rows], as.numeric(pd))$F1
    })

@@

  gp = ggplot(df, aes(y=mean,x=lambda.at)) + geom_point() + 
          geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme)) + labs(y='mean of confusion matrix (P)') +
          geom_point(data=subset(df, mean == max(mean)), aes(y=mean, x=lambda.at, colour="max"), size=3 ) +
          geom_point(data=df[which.max(df$mean)+1,], aes(y=mean, x=lambda.at, colour="max+1"), size=3) +
          geom_point(data=df[which.max(df$mean)-1,], aes(y=mean, x=lambda.at, colour="max-1"), size=3) +
          geom_hline(yintercept=df[which.max(df$mean)-1,'mean'], colour='grey') +
          annotate("text", x=5, y=max(df$mean)+0.0005, label=round(df[which.max(df$mean)-1,'mean'],2)) +
          theme(axis.title.x = element_text(size = 15, vjust=-.2)) + scale_colour_discrete(name = "Best lambda values") +
          labs(title="CV Performance")

  return(list('cv.class'=cv.class, 'f1'=auc[which.max(df$mean)-1], 'lambda.opt'=lambda.opt, 'lambdas'=df, 'plot'=gp, 'non.zero.cf'=nzcf))
}

@@

grid.arrange(cv.patient$plot)

## check how often the feature(s) is selected at that lamda in each split. "stability selection"

for (i in 1:length(cv.patient$non.zero.cf)) {
  df = as.data.frame(cv.patient$non.zero.cf[[i]])
  colnames(df) = i
  coef.opt = merge(coef.opt, df, by='row.names', all.x=T)
  rownames(coef.opt) = coef.opt$Row.names
  coef.opt$Row.names = NULL
  coef.opt[,(i+1)] = as.integer(!is.na(coef.opt[,(i+1)]))
}
coef.opt = coef.opt[order(sapply(rownames(coef.opt), function(x) as.numeric(unlist(strsplit(x, ':'))[1])  )),]

@@

@@

cv.nohgd = cv.patient.glmnet(x=dysplasia.df[samples,], y=labels[samples], lambda, subset(patients, Samplename %in% samples))
#lambda.opt = cv.patient$lambda.opt
grid.arrange(cv.nohgd$plot + ggtitle("CV Performance, No HGD/IMC samples"))
#coef.opt = as.data.frame(non.zero.coef(fit0, lambda.opt))

samples = intersect(rownames(dysplasia.df), subset(info, Pathology %nin% c('HGD', 'IMC', 'LGD'))$Samplename)
cv.nolgd = cv.patient.glmnet(x=dysplasia.df[samples,], y=labels[samples], lambda, subset(patients, Samplename %in% samples))
#lambda.opt = cv.patient$lambda.opt
grid.arrange(cv.nolgd$plot + ggtitle("CV Performance, No HGD/IMC/LGD samples"))

get.next.samples<-function(patient.data, i) {
  as.vector(unlist(sapply(patient.data, function(pd) {
    j = i
    pd$info = arrange(pd$info, Endoscopy.Year, Pathology)

@@

nonps = get.next.samples(patient.data[subset(sum.patient.data, Status == 'NP')$Patient], 0)
plots = list()
for (i in seq(1,35,5)) {
  ps = get.next.samples(patient.data[subset(sum.patient.data, Status == 'P')$Patient], i)
  samples = c(nonps, ps)
  cv.back = cv.patient.glmnet(x=dysplasia.df[samples,], y=labels[samples], lambda, subset(patients, Samplename %in% samples))
  plots[[ as.character(i) ]] = cv.back$plot + ggtitle(paste("CV Performance - ", length(ps), "progressor samples"))
}

do.call('grid.arrange', c(plots, ncol=2, top="Peeling back timepoints in progressors"))

>>>>

sum.patient.data = summarise.patient.info(patient.info)
sum.patient.data = as.data.frame(subset(sum.patient.data, Patient %in% names(patient.data))) ## For now
pander(sum.patient.data[,c('Patient', 'Status', 'start.year', 'end.year','total.samples','total.endos','highest.path')])

@@

```{r xvalpt, message=F, warning=F, echo=F, fig.height=6, fig.width=6}
precisionRecall<-function(actual, predicted) { ## needs to calculate for both 0 and 1
    retrieved <- sum(predicted==1)
    recall <- sum(predicted==1 & actual==1) / sum(actual==1)
    sum(predicted==0 & actual==0) / sum(actual==0)

    precision <- sum(predicted==1 & actual==1) / retrieved
    f1 = signif(2*(1/(1/precision+1/recall)), 2)
    return(data.frame('precision'=precision, 'recall'=recall, 'F1'=signif(f1, 2)))
}

@@

cv.patient.glmnet<-function(x,y,lambda,pts) {
  message(paste("Running", length(unique(pts$group)), "folds"))

  cv.class = matrix(nrow=5, ncol=length(lambda))
  f1 = matrix(nrow=5, ncol=length(lambda))
  prec = f1; recall = f1

  fit.e = list()
  for (i in 1:length(unique(patients$group))) {  ## This is not stable. 
    message(paste(i, "fold"))
    test.rows = which(rownames(x) %in% subset(pts, group == i)$Samplename)
    test = x[test.rows,]
    training = x[-test.rows,]
    # pre-spec lambda seq
    fit <- glmnet(training, y[-test.rows], lambda=lambda, family='binomial', alpha=1) 
    # autoplot(fit) + theme(legend.position="none")

    pred.class <- predict(fit, test, type='class') # check this later

    proc = do.call(rbind, apply(pred.class, 2, function(pd) {
      if (sum(as.numeric(pd)) <= 0) return(0)
      #as.numeric(pROC::auc(pROC::roc(response=as.numeric(pd), predictor=labels[test.rows])))  
      precisionRecall(labels[test.rows], as.numeric(pd))
    }))

    f1[i,] = proc$F1
    prec[i,] = proc$precision
    recall[i,] = proc$recall

@@

  gp = ggplot(df, aes(y=mean,x=lambda.at)) + geom_point() + 
          geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme)) + labs(y='mean of confusion matrix (P)') +
          geom_point(data=subset(df, mean == max(mean)), aes(y=mean, x=lambda.at, colour="max"), size=3 ) +
          geom_point(data=df[which.max(df$mean)+1,], aes(y=mean, x=lambda.at, colour="max+1"), size=3) +
          geom_point(data=df[which.max(df$mean)-1,], aes(y=mean, x=lambda.at, colour="max-1"), size=3) +
          geom_hline(yintercept=df[which.max(df$mean)-1,'mean'], colour='grey') +
          annotate("text", x=5, y=max(df$mean)+0.005, label=round(df[which.max(df$mean)-1,'mean'],2)) +
          annotate("text", x=6, y=max(df$mean)-0.005, label=round(median(df$sme),4)) +
          theme(axis.title.x = element_text(size = 15, vjust=-.2)) + scale_colour_discrete(name = "Best lambda values") +
          labs(title="CV Performance")

  return(list('max.cm'=df[which.max(df$mean)-1,'mean'], 'f1'=mean(f1[,which.max(df$mean)-1]), 'precision'=mean(prec[,which.max(df$mean)-1]), 'recall'=mean(recall[,which.max(df$mean)-1]),
              'lambda.opt'=lambda.opt, 'lambdas'=df, 'plot'=gp, 'non.zero.cf'=nzcf))
}

@@

grid.arrange(cv.patient$plot + ggtitle(paste("CV Performance, 5-fold xval. mean f1 at opt lambda:", cv.patient$f1)))

## check how often the feature(s) is selected at that lamda in each split. "stability selection"
coef.stability<-function(opt, nz.list) {
  for (i in 1:length(nz.list)) {
    df = as.data.frame(nz.list[[i]])
    colnames(df) = i
    opt = merge(opt, df, by='row.names', all.x=T)
    rownames(opt) = opt$Row.names
    opt$Row.names = NULL
    opt[,(i+1)] = as.integer(!is.na(opt[,(i+1)]))
  }
  opt = opt[order(sapply(rownames(opt), function(x) as.numeric(unlist(strsplit(x, ':'))[1])  )),]
  return(opt)
}

coef.opt.stability = coef.stability(coef.opt, cv.patient$non.zero.cf)

@@

### Feature stability

`r length(which(rowSums(coef.opt.stability[,2:6])>= 3))` features are stable in 3:5 splits

`r length(which(rowSums(coef.opt.stability[,2:6])>= 4))` features are stable in 4:5 splits

`r length(which(rowSums(coef.opt.stability[,2:6])>= 5))` features are stable in 5:5 splits

Optimal model?
```{r}

gfit = glm(labels~.,data=cbind.data.frame(dysplasia.df[,names(non.zero.coef(fit0, lambda.opt))], labels), family=binomial(link='logit'))
autoplot(gfit)

@@

cv.nohgd = cv.patient.glmnet(x=dysplasia.df[samples,], y=labels[samples], lambda, subset(patients, Samplename %in% samples))
#lambda.opt = cv.patient$lambda.opt
grid.arrange(cv.nohgd$plot + ggtitle(paste("CV Performance, No HGD/IMC samples. mean f1:", cv.nohgd$f1)))
nohgd.stab = coef.stability(coef.opt, cv.nohgd$non.zero.cf)

samples = intersect(rownames(dysplasia.df), subset(info, Pathology %nin% c('HGD', 'IMC', 'LGD'))$Samplename)
cv.nolgd = cv.patient.glmnet(x=dysplasia.df[samples,], y=labels[samples], lambda, subset(patients, Samplename %in% samples))
grid.arrange(cv.nolgd$plot + ggtitle(paste("CV Performance, No HGD/IMC/LGD samples. mean f1:", cv.nohgd$f1)))
nolgd.stab = coef.stability(coef.opt, cv.nolgd$non.zero.cf)

#### No HGD feature stability

`r length(which(rowSums(nohgd.stab[,2:6])>= 3))` features are stable in 3:5 splits

`r length(which(rowSums(nohgd.stab[,2:6])>= 4))` features are stable in 4:5 splits

`r length(which(rowSums(nohgd.stab[,2:6])>= 5))` features are stable in 5:5 splits

#### No LGD feature stability

`r length(which(rowSums(nolgd.stab[,2:6])>= 3))` features are stable in 3:5 splits

`r length(which(rowSums(nolgd.stab[,2:6])>= 4))` features are stable in 4:5 splits

`r length(which(rowSums(nolgd.stab[,2:6])>= 5))` features are stable in 5:5 splits

### Leave one patient out

```{r loop, echo=F, message=F, warning=F, fig.height=40, fig.width=30}
fpr = data.frame(matrix(ncol=5,nrow=0))
# Randomize patients
pt.samp = sample(sum.patient.data$Patient)

#plots = list()

for (pt in pt.samp) {
  samples = as.vector(unlist(lapply(patient.data[subset(sum.patient.data, Patient != pt)$Patient], function(pt) pt$info$Samplename)))

  cv.back = cv.patient.glmnet(x=dysplasia.df[samples,], y=labels[samples], lambda, subset(patients, Samplename %in% samples))
#  print(median(cv.back$lambdas$sme))
 # print( paste(cv.back$f1, cv.back$precision) )
   fpr = rbind(fpr, c(cv.back$max.cm, cv.back$f1, cv.back$precision, cv.back$recall, length(samples)))

#  plots[[ pt ]] = cv.back$plot + ggtitle(paste("# Samples:", length(samples))) + theme(legend.position = "none")
}
colnames(fpr) = c('performance','f1','prec','recall','n.samples')

ggplot(fpr, aes(performance, f1)) + 
  geom_point() + 
  ylim(0,1) + xlim(0,1) + labs(title="LOO (patient)", x='CV Performance', y='f1 statistic')

## Peel back timepoints

```{r, echo=F, message=F, warning=F}
get.next.samples<-function(patient.data, i) {
  as.vector(unlist(sapply(patient.data, function(pd) {
    j = i
    pd$info = arrange(pd$info, Endoscopy.Year, Pathology)

@@

#subset(sum.patient.data, Status == 'P' & total.endos >= 8)

#cv.back = cv.patient.glmnet(x=dysplasia.df[samples,], y=labels[samples], lambda, subset(patients, Samplename %in% samples))
#print(cv.back$f1)

nonps = get.next.samples(patient.data[subset(sum.patient.data, Status == 'NP')$Patient], 0)
#plots = list()
fpr = data.frame(matrix(ncol=4,nrow=0))
for (i in c(0,1,seq(2,25,3))) {
  ps = get.next.samples(patient.data[subset(sum.patient.data, Status == 'P')$Patient], i)
  samples = get.next.samples(patient.data, i)
  #samples = c(nonps, ps)
  cv.back = cv.patient.glmnet(x=dysplasia.df[samples,], y=labels[samples], lambda, subset(patients, Samplename %in% samples))
  print( paste(cv.back$f1, cv.back$precision) )
  fpr = rbind(fpr, c(cv.back$f1, cv.back$precision, cv.back$recall, length(ps)))
#  plots[[ as.character(i) ]] = cv.back$plot + ggtitle(paste(length(ps), "progressor samples. mean f1:", cv.back$f1))
}
colnames(fpr) = c('f1','prec','recall','n.samples')
pander(fpr[1:5,], justify='left')
#do.call('grid.arrange', c(plots, ncol=2, top="Peeling back timepoints in progressors"))

ps = get.next.samples(patient.data[subset(sum.patient.data, Status == 'P')$Patient], 11)
samples = c(nonps, ps)
  cv.back = cv.patient.glmnet(x=dysplasia.df[samples,], y=labels[samples], lambda, subset(patients, Samplename %in% samples))
  print(median(cv.back$lambdas$sme))
  cv.back$plot + ggtitle(paste("CV Performance - ", length(ps), "progressor samples"))

info = subset(patient.info, Samplename %in% samples)
head(info)

pander(summarise.patient.info(subset(info, Status == 'P'))[,c(1:7)], justify='left')

coef.opt.stability = coef.stability(coef.opt, cv.back$non.zero.cf)

pander(coef.opt.stability[which(rowSums(coef.opt.stability[,2:6]) > 0),], justify='left', caption="Features initially identified that also are found in the subset of samples")
